Diagnosing Child Poverty Reduction Interventions

### 2024-09-03 minos/outcomes/CPR_Intervention_Diagnostics.Rmd
Added plots to show hh_income by quintile with density ridge plots
Additional changes forgot to save

@@

@@

@@

xgb.ggplot.shap.summary(sf12)
```

>>>>

```{r}
colnames(base.dat)
```

@@

Firstly, can't rely on the income_quintile variable as this is a variable created in start year which does not transition. Therefore we will need to calculate an income quintile variable for both dataframes.

```{r}
base.dat <- base.dat %>%
  group_by(time) %>%
  mutate(income_quint = ntile(hh_income, 5))

int.dat <- int.dat %>%
  group_by(time) %>%
  mutate(income_quint = ntile(hh_income, 5))
```

@@

density_plot <- function(base, int) {
  p <- ggplot(base, aes(x = hh_income, y = time, group = time)) +
    geom_density_ridges(color = 'blue', alpha = 0.3) +
    geom_density_ridges(data = int, aes(color = 'red', alpha = 0.3)) +
    theme(legend.position = 'none') +
    labs(title = 'Household Income Density Plots')

  return(p)
}

p <- density_plot(base.dat, int.dat)

p <- p + facet_wrap(~ income_quint, ncol=1)

print(p)

# ggplot(base.dat, aes(x = hh_income, y = time, group = time)) +
#   geom_density_ridges(color = 'blue', alpha = 0.3) +
#   geom_density_ridges(data = int.dat, aes(color = 'red', alpha = 0.3)) +
#   theme(legend.position = 'none') +
#   labs(title = 'Household Income Density Plots') +
#   facet_wrap(~ income_quint, ncol=1)
```

```{r}
density_plot_quintile <- function(base, int, quintile) {

  base <- base %>%
    filter(income_quint == quintile) %>%
    select(pidp, time, hh_income) %>%
    rename(Baseline = hh_income)
  int <- int %>%
    filter(income_quint == quintile) %>%
    select(pidp, time, hh_income) %>%
    rename(Intervention = hh_income)

  combined <- merge(base, int, by = c('pidp', 'time'))
  combined <- combined %>%
    pivot_longer(cols = Baseline:Intervention,
                 names_to = 'Source',
                 values_to = 'hh_income')

  p <- ggplot(combined, aes(x = hh_income, y = time, group = interaction(time, Source), color = Source)) +
    geom_density_ridges() +
    labs(title = 'Household Income Density Plots',
         subtitle = paste0('Quintile ', quintile))

  return(p)
}

density_plot_quintile(base.dat, int.dat, quintile = 1)
density_plot_quintile(base.dat, int.dat, quintile = 2)
density_plot_quintile(base.dat, int.dat, quintile = 3)
density_plot_quintile(base.dat, int.dat, quintile = 4)
density_plot_quintile(base.dat, int.dat, quintile = 5)
```
```{r}
density_plot_quintile_2 <- function(base, int, quintile) {
  base <- base %>%
    filter(income_quint == quintile) %>%
    select(pidp, time, hh_income) %>%
    rename(Baseline = hh_income)
  int <- int %>%
    select(pidp, time, hh_income) %>%
    rename(Intervention = hh_income)

  combined <- merge(base, int, by = c('pidp', 'time'))
  combined <- combined %>%
    pivot_longer(cols = Baseline:Intervention,
                 names_to = 'Source',
                 values_to = 'hh_income')

  # Calculate the 1st and 99th percentiles
  lower_cutoff <- quantile(combined$hh_income, 0.01)
  upper_cutoff <- quantile(combined$hh_income, 0.99)

  filtered_combined <- combined %>%
    filter(hh_income > lower_cutoff, hh_income < upper_cutoff)

  p <- ggplot(filtered_combined, aes(x = hh_income, y = time, group = interaction(time, Source), color = Source)) +
    geom_density_ridges() +
    labs(title = 'Household Income Density Plots',
         subtitle = paste0('Quintile ', quintile))
  return(p)
}

density_plot_quintile_2(base.dat, int.dat, quintile = 1)
density_plot_quintile_2(base.dat, int.dat, quintile = 2)
density_plot_quintile_2(base.dat, int.dat, quintile = 3)
density_plot_quintile_2(base.dat, int.dat, quintile = 4)
density_plot_quintile_2(base.dat, int.dat, quintile = 5)
```

@@

sf12$params
importance_matrix <- xgb.importance(model = sf12)
print(importance_matrix)

xgb.plot.importance(importance_matrix)
```
